---
title: "Coral Recruitment Patterns in Different Reef Environments"
author: "Hannah Merges"
date: "`r Sys.Date()`"
output: powerpoint_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
```

## What is a coral? 
- Plant (photosynthetic algal symbiont)
- Animal (coral host) 
- Rock (calcium carbonate skeleton)

## What is coral recruitment? 
- The introduction of new individuals to a community 
- Increases genetic diversity 
- Aids in post-disturbance recovery 

## Different types of reef environments 
- Fringing reef (nearest the shore)
- Backreef (close to where the waves break)
- Forereef (usually closest to the reef wall which is in in open water and slopes/drops off to larger depths)

## Loading libraries 

```{r, echo = TRUE}
library(tidyverse)
library(here)
library(janitor)
library(PNWColors)
```

## Read in the data 

```{r }
recruits <- read_csv(here("Data","coral_recruit_tile_spat_counts_2006-2016_20180626.csv"))
```

## Tidy the data 
In a way that is useful for the plots and output we are trying to create 
Want to look at the change in recruit density over the years along the western shore particuarly the lagoon 

```{r}
## ideally looking to combine total counts of all tiles and season into one count per year by species 
recruityears <- recruits %>% 
  filter(habitat=="Lagoon") 
  
```

