---
title: "Coral Recruitment Patterns in Different Reef Environments"
author: "Hannah Merges"
date: "`r Sys.Date()`"
output: powerpoint_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE, 
                      fig.path = ".../Outputs/")
```

## What is a coral? 
- Plant (photosynthetic algal symbiont)
- Animal (coral host) 
- Rock (calcium carbonate skeleton)

## What is coral recruitment? 
- The introduction of new individuals to a community 
- Increases genetic diversity 
- Aids in post-disturbance recovery 

## Different types of reef environments 
- Fringing reef (nearest the shore)
- Backreef (close to where the waves break)
- Forereef (usually closest to the reef wall which is in in open water and slopes/drops off to larger depths)

## Loading libraries 

```{r, echo = TRUE}
library(tidyverse)
library(here)
library(janitor)
library(PNWColors)
```

## Read in the data 

```{r }
recruits <- read_csv(here("Data","coral_recruit_tile_spat_counts_2006-2016_20180626.csv"))
```

## Tidy the data 
In a way that is useful for the plots and output we are trying to create 
Want to look at the change in recruit density over the years along the western shore particularly the lagoon 

```{r}
## ideally looking to combine total counts of all tiles and season into one count per year by species 

recruityears <- recruits %>%
  pivot_wider(names_from = side, values_from = count) %>% 
  filter(family!="Unidentified", ##this filters out these families and says side does NOT include"A"
         family!="Other") %>% ## was going to filter out by shore, but "west" is only in 2006 
  select(-location, -A) %>%
  mutate(total_count = B+S+T) %>% 
  group_by(nominal_year, family, habitat) %>% 
  summarize(sum_total=sum(total_count))
  
```

## Plot this Data 
Shows the total number of recruits by family per year. 

```{r, recruit_total}
ggplot(data=recruityears, 
       aes(x=nominal_year, 
           y=sum_total, 
           fill=family)) + 
  geom_col() + 
  labs(x="Year", 
       y="Total Number of Recruits", 
       title="Total Number of Recruits by Family from 2006-2016", 
       fill="Coral Family") + 
  facet_wrap(~habitat) + 
  theme_linedraw() + 
  theme(plot.title=element_text(hjust=0.5),
        axis.text.x=element_text(size=10, angle = 45, vjust=0.5), ##to get the x axis slightly angled so text is not on top of each other. vjust moves text away from the actual axis 
        axis.text.y=element_text(size=10), 
        axis.title.x=element_text(size=12),
        axis.title.y=element_text(size=12)) +
  scale_x_continuous(breaks = c(2006, 2007, 2008, 2009, 2010, 2011, 2012, 2013, 2014, 2015, 2016)) + ##allows you to specifically decide what you want the column breaks to be 
  scale_fill_manual(values=pnw_palette("Sunset2", n=3))
ggsave(here("Outputs", "totalrecruitment2.jpg"), 
       width=10, height=7)

  
```

## Now determine which side of the tile most recruits settle on 
```{r}
tileside <- recruits %>% 
  filter(family!="Unidentified", ##this filters out these families 
         family!="Other", 
         side!="A") %>% 
  select(-location) %>% ##removes columns by name
  group_by(family, side) %>% 
  summarize(sum_total=sum(count))

```

## plot it 
```{r}
ggplot(data=tileside,
      aes(x=side, 
          y=sum_total, 
          fill=side)) + 
  geom_col() + 
  facet_wrap(~family) + 
  theme_bw() + 
  labs(y="Total Number of Recruits", 
       x="Side of Recruitment Tile", 
       title="Corals Settle More on the Bottom of Recruitment Tiles",
       fill="Side of Tile") + 
  theme(plot.title=element_text(hjust=0.5),
        axis.text.x=element_text(size=10), ##to get the x axis slightly angled so text is not on top of each other. vjust moves text away from the actual axis 
        axis.text.y=element_text(size=10), 
        axis.title.x=element_text(size=12),
        axis.title.y=element_text(size=12)) + 
  scale_x_discrete(labels = c("Bottom", "Side", "Top")) + 
  scale_fill_manual(values = pnw_palette("Bay", n=3))
ggsave(here("Outputs","tilelocation.jpg"), 
       height=7, width=10)
```


